Apparent preferences for cognitive effort fade when multiple forms of effort

### 2020-11-13 Cost_manuscript.Rmd
tweking citations

---
title: Apparent preferences for cognitive effort fade when multiple forms of effort
  and delay are interleaved in a foraging environment
author: "Claudio Toro-Serey, Gary Kane, & Joseph T. McGuire"
header-includes:
- \usepackage{setspace}\doublespacing
- \usepackage[left]{lineno}
- \linenumbers
- \usepackage{float}
- \newcommand{\beginsupplement}{ \setcounter{table}{0} \renewcommand{\thetable}{S\arabic{table}}
  \setcounter{figure}{0} \renewcommand{\thefigure}{S\arabic{figure}} }
- \usepackage{indentfirst}
output:
  pdf_document:
    highlight: haddock
    keep_tex: no
  word_document: #default
    reference_docx: Style_reference_draft.docx
  html_document:
    df_print: paged
csl: apa.csl
indent: true
#subtitle: 'Abbreviated Title: Heterogeneity in the functional topography of mPFC'
bibliography: cost_paper.bib
---

@@

Department of Psychological and Brain Sciences & Center for Systems Neuroscience, Boston University, Boston, USA

@@

  Evaluating whether to pursue a desired outcome often requires assessing the demands required for obtaining it. In psychology and economics, demands for both time and effort have traditionally been identified as costs that detract from the net subjective value of rewards [@Westbrook2015; @Shenhav2017]. Indeed, decision makers tend to prefer low-effort and immediate rewards over larger rewards that demand longer delays and higher effort [@Ainslie1975; @Hull1943; @Kool2018; @Kool2010; @Treadway2009; @Westbrook2013]. However, people often change their attitudes towards exerting effort in everyday decisions, resulting in pursual of effortful avenues instead. Emerging experimental and theoretical work has chiefly explained this phenomenon by suggesting a reciprocal relationship between effort and outcome, whereby exerting effort boosts the perceived value of its outcomes [@Kacelnik2002; @Mochon2012; @HernandezLallement2014], and positive outcomes imbue value into their preceding effortful behavior [i.e. learned industriousness; @Eisenberger1992]. Though intuitively valid and empirically supported, this explanation leaves open questions about what features of an effortful task can make it attractive in its own right, and how contextual factors mold its perceived costs [@Inzlicht2018]. Understanding these intrinsic qualities could improve our ability to motivate completion of effortful daily activities, such as schoolwork or physical exercise.

  What is often missing from existing accounts is the common case of choosing among types of behavioral demands. Studies have traditionally compared the reduction of subjective value imposed by each type of demand, often through discounting functions that capture how much less appealing a reward becomes as a function of the amount or intensity of its associated behavior [@Ainslie1975; @Frederick2002; @Kable2007; @Kool2010; @Westbrook2013; @Green1994). Although this line of work suggests that discounting functions are similar for delay and effort [@Prevost2010; @Massar2015; @Seaman2018], there is ongoing debate about the shape of the discounting function across effort types [@Arulpragasam2018; @Prevost2010; @Chong2016; @Klein-Flugge2016; @Kool2018], and discount rates do not always correlate across cost domains [@Westbrook2013; @Seaman2018]. However, sporadic preferences for effort could arise from the availability of other types of demands in the same choice environment—in other words, from its context [@Kool2018]. For example, exerting high effort for a reward might be aversive, but the same level of work might become valuable when compared to the boredom of passive waiting (e.g. taking the cumbersome way home rather than waiting for a delayed train, even if the resulting time of arrival is the same). Conversely, even though people tend to donate more to charitable events that require them to perform effortful activities (i.e. the martyrdom effect), they are also less likely to engage in such events once an easier alternative to raise funds becomes available [@Olivola2013]. Recent work has proposed that the subjective costs produced by delays and cognitive effort is modulated by the possibility of missing out on better concurrent alternatives, namely the opportunity cost of time [OC; @Fawcett2012; @Kurzban2013; @Otto2019]. This evidence sets the stage for examining what people would actually select when given the option for multiple types of demands, including the unexplored question of deciding between cognitive and physical effort [@Inzlicht2018; @Kool2018].

  Single choice foraging paradigms provide a natural way to probe preferences among demands, by asking individuals whether the current offer outweighs what could be obtained and experienced during that time instead [@Stephens2019]. Foraging has recently attracted high scientific interest due to its ecological validity, which is rooted in evolutionary choice behaviors [@Hayden2018; @Mobbs2018], and offers a number of desirable qualities. First, it allows for the estimation of optimal, reward maximizing choice patterns based on the opportunity cost of time [@Charnov1975; @Krebs1977]. We can therefore interpret observed deviations in acceptance rates from such optimal behavior in addition to comparing them among demands. Second, the sequential nature of foraging allows individuals to emphasize different elements of the decision context. For example, the cost (or value) of effort or delay could globally depend on the availability of alternative demands (as suggested above), impacting the perceived opportunity cost of time. Alternatively, it could be intrinsic and robust to such manipulation, and/or depend on task-specific factors that are adjusted once individuals experience different forms of behavioral costs. Foraging rests on an abundant modeling history that can help to computationally capture such emphases [@Stephens2019].

@@

  All experimental tasks were implemented using PsychoPy 2 (v1.85.1, Pierce2019) on a Macbook Pro laptop. In this task, participants foraged for monetary rewards in an environment in which each trial required either physical effort or cognitive effort for a set period of time (the “handling time”), or an equivalent unfilled delay (Figure 1). Their goal was to maximize their gains within a fixed amount of time. On each trial a monetary offer was displayed for 2 s, and participants had the opportunity to expend time and/or effort during the handling time in order to earn it. Upon completion of a trial participants saw a 2 s window displaying the reward obtained, which was followed by a travel time to the next offer. Alternatively, the participant could quit at any point during the handling time by pressing the spacebar on the computer, and immediately start traveling.

  Each participant was assigned to one of three types of costs (cognitive effort, physical effort, or delay), and a fourth group of equal size faced an effortless physical task that involved minimal gripping. Each group was unaware that other cost conditions existed. Participants exerted physical effort by maintaining grip on a handheld dynamometer (XXXX, BIOPAC Systems, United States) using their dominant hand. Gripping requirements were calibrated at 20% of maximum voluntary contraction (MVC, acquired at the beginning of the session). Cognitive effort entailed switching among Stroop, dot motion coherence, and flanker tasks. In Stroop, one of three color names was displayed on the screen (red, blue or green), with a font color that was either congruent (e.g. word red painted in red) or incongruent (e.g. word red painted in blue). Participants had to select the color of the font, not the word displayed (i.e. they had to suppress their tendency to read the word). For motion coherence, 100 solid white dots moved on the screen. A fraction of these dots moved cohesively to the left or right, while the rest moved in random directions (coherence could be either 30% or 40%, uniformly sampled). Participants had to respond with the direction of the cohesive set of dots. In flanker, rows of arrowheads pointed either to the left or the right (maximum of 3 rows, 3 to 13 arrowheads per row). Participants responded with the direction of the center arrowhead, which could point in the same or opposite direction from its neighbors. These tasks were configured so that responses always involved a left or right key press (e.g. for Stroop, two colored circles were presented at each side of the screen). During the handling time, cognitive tasks and their configurations were randomly sampled, and were presented for 1 s followed by a 1 s inter-stimulus interval. Participants were asked to respond within each task’s presentation time. Before the experiment, participants trained in each cognitive task until they correctly performed six consecutive tasks of each kind. While in the handling time, if participants failed to maintain above-threshold gripping or made two mistakes during the cognitive task, they were forced to travel and missed the reward.

  There were three block types, in which handling times of 2, 10, or 14 seconds were paired with travel times of 14, 6, and 2 seconds, respectively (note that all combinations add up to 16 seconds). Timing parameters were held constant within each 7-min block. Each of these blocks was experienced in a pseudo-randomized order, and repeated in the same order after a short break in order to probe choice stability (total session length = 42 minutes). Reward amounts varied uniformly per trial (4, 8, or 20 cents), with the constraint that every reward was presented twice every six trials. This prevented sequences from being dominated by a single amount during any window of time. Timing information was disclosed at the beginning of each block, and rewards displayed during a 2 s offer window before each trial began. Participants received training prior to the experimental session, and were told about all possible environmental statistics in order to preclude biases due to experience-dependent learning (Dundon2020; Garrett2020).

@@

  Foraging theory posits that accepting a delayed reward should depend on the opportunity cost (OC) of time incurred in obtaining it, given by the richness of the environment (Stephens2019). In this study, the richness of the environment was manipulated by the length of the handling and travel times, where shorter travel times produced richer environments. Since time combinations were fixed per block, we calculated each block’s optimal accept/reject strategy by computing all decision strategies according to the following equation:

@@

  All analyses were performed in R 3.5.2 (RTeam2018). First we tested whether decision makers integrated delay and reward information. To address the prediction that participants would be more likely to accept higher reward and shorter handling time trials, we fit a mixed-effects logistic regression to predict trial-wise acceptances (using the lme4 package, Bates2015), giving a random intercept to each subject to account for biases in participant behavior. We included regressors for handling time and reward, as well as two covariates that probed the influence of recent history on choices. The first was an autoregressive regressor containing the number of consecutive offer misses prior to a given trial (misses could be due to quits or forced travels). The second regressor tracked the sum of the previous *n* offers. We identified the *n* that minimized Akaike's Information Criterion (AIC) among 6 versions of the model (tracing from 2 previous trials up to 7, beyond which the model failed to converge), and reported all coefficients from the winning model.

  Our foraging task was configured such that over and under accepting were detrimental to total earnings. To confirm this, we fit a general linear model with constant, linear, and quadratic terms to estimate the correspondence between proportion of trials completed (independent variable) and total earnings (dependent variable). A significant quadratic coefficient thus would signal that the task statistics operated as expected. Next, to determine the optimality of each group’s decisions, we performed two-sided one-sample t-tests to see if the proportion of acceptances for each time/reward combination was significantly different from the optimal rate (see Operationalization of Cost). This resulted in 36 independent tests (3 rewards amounts, 3 handling/travel time combinations, and 4 groups), so we corrected for multiple comparisons using False Discovery Rate (FDR, Benjamini2001).

@@

  We started by fitting a base model derived from the normative predictions of the Marginal Value Theorem (MVT; Charnov, 1975), which states that a delayed reward should be accepted if its earning rate surpasses the opportunity cost (OC) of time incurred in obtaining it. This decision rule can be expressed as

@@

$$\begin{aligned}
P(accept)_i = \dfrac {1} {1 + e^{-1/\beta(Model_{base})}}
\end{aligned}$$

@@

  A final model considered the relationship between effortful engagement and the perceived length of time. Being immersed in a task can make time transpire subjectively faster [Csikszentmihalyi2004], which might have made the cognitive task attractive. Time sensitivity has been previously used to improve fits on classic hyperbolic models of delay discounting [Mckerchar2009]. It is possible that such subjective handling time was recalibrated in the within-subjects experiment, resulting in the observed convergance of acceptances across demands. We therefore added a time sensitivity parameter $S$ to the model, with a separate value associated with each type of demand. The values of $S_{demand}$ could then be recalibrated as participants experienced them, averaged according to the following formula [adapted from @Sutton2018]

@@

  in which high values of the free parmeter $\alpha_w$ denoted that decision makers incorporated the recalibrated subjective time early on in the experiment. Similar to McKerchar et al. (2009), the values of $mS_i$ exponentially modulated every instance of the handling time ($H_i$) from the adaptive model. In this way, subjective time perception produced by different demands could influence choice in two ways: retrospectively affecting the value of $\gamma$ through the perceived length of recent trials

@@

  The next set of analyses examined the hypothesis that effort group would uniformly accept more trials. Figure 3 (left) shows that the cognitive effort group consistently accepted more offers, which resulted in lower earnings (middle). One-way ANOVAs showed that differences among groups were significant both for overall acceptance rates (F(`r propAov[[1]]$Df[1]`, `r propAov[[1]]$Df[2]`) = `r round(propAov[[1]][1, 4], digits = 2)`, *p* `r report_p(propAov[[1]][1, 5])`), and total earnings (F(`r earnAov[[1]]$Df[1]`, `r earnAov[[1]]$Df[2]`) = `r round(earnAov[[1]][1, 4], digits = 2)`, *p* `r report_p(earnAov[[1]][1, 5])`). Post-hoc permutations (5000 iterations) comparing mean acceptance rates between every pair of costs confirmed that acceptance rates of the cognitive effort group (mean = `r filter(overallSumms, Cost == "Cognitive")$mProp`, SD = `r filter(overallSumms, Cost == "Cognitive")$sdProp`) were higher than those of the physical (mean = `r filter(overallSumms, Cost == "Physical")$mProp`, SD = `r filter(overallSumms, Cost == "Physical")$sdProp`; *p* `r report_p(filter(propTests, contrast == "Physical < Cognitive")$permPval)`, Cohen's D = `r filter(propTests, contrast == "Physical < Cognitive")$cohenD`) and wait groups (mean = `r filter(overallSumms, Cost == "Wait")$mProp`, SD = `r filter(overallSumms, Cost == "Wait")$sdProp`; *p* `r report_p(filter(propTests, contrast == "Cognitive > Wait")$permPval)`, Cohen's D = `r filter(propTests, contrast == "Cognitive > Wait")$cohenD`). It also showed that those faced with the easy task accepted more than those in the wait group, although responses from the easy group were more variable (mean Easy = `r filter(overallSumms, Cost == "Easy")$mProp`, SD = `r filter(overallSumms, Cost == "Easy")$sdProp`; *p* `r report_p(filter(propTests, contrast == "Wait < Easy")$permPval)`, Cohen's D = `r filter(propTests, contrast == "Wait < Easy")$cohenD`) (all other *p* > 0.05). The higher acceptance rates produced by the cognitive effort group resulted in significantly lower earnings (mean = `r filter(overallSumms, Cost == "Cognitive")$mEarn`, SD = `r filter(overallSumms, Cost == "Cognitive")$sdEarn`) than both physical (mean = `r filter(overallSumms, Cost == "Physical")$mEarn`, SD = `r filter(overallSumms, Cost == "Physical")$sdEarn`; *p* `r report_p(filter(earnTests, contrast == "Physical > Cognitive")$permPval)`, Cohen's D = `r filter(earnTests, contrast == "Physical > Cognitive")$cohenD`) and wait groups (mean = `r filter(overallSumms, Cost == "Wait")$mEarn`, SD = `r filter(overallSumms, Cost == "Wait")$sdEarn`; *p* `r report_p(filter(earnTests, contrast == "Cognitive < Wait")$permPval)`, Cohen's D = `r filter(earnTests, contrast == "Cognitive < Wait")$cohenD`). In addition, we found that the wait group earned more than the physical (*p* `r report_p(filter(earnTests, contrast == "Physical < Wait")$permPval)`, Cohen's D = `r filter(earnTests, contrast == "Physical < Wait")$cohenD`) and easy groups (mean Easy = `r filter(overallSumms, Cost == "Easy")$mEarn`, SD = `r filter(overallSumms, Cost == "Easy")$sdEarn`; *p* `r report_p(filter(earnTests, contrast == "Wait > Easy")$permPval)`, Cohen's D = `r filter(earnTests, contrast == "Wait > Easy")$cohenD`). The proportion of forced travels for cognitive effort was `r round(propFails[2, 2], digits = 2)` across participants, and 0 for either physical or easy conditions, making performance an unlikely explanation for the observed differences in choice.

@@

  Stability was defined as a persistence in acceptance patterns throughout the experimental session. The scatter plot on Figure 5 (bottom-left) shows the overall proportion accepted by each participant before and after the mid-point break. Paired permutations showed no differences in acceptances for wait and easy groups (*p* > 0.05), but we found significantly lower acceptances for physical (*p* `r report_p(prepost_btw$Perms[1])`, Cohen's D = `r round(prepost_btw$CohenD[1], digits = 2)`) and cognitive (*p* `r report_p(prepost_btw$Perms[2])`, Cohen's D = `r round(prepost_btw$CohenD[2], digits = 2)`) groups. Figure 6 (left) shows that the reduction acceptances in effort groups over time was steeper for longer handling times, a hallmark of effortful demands (Treadway et al, 2009). We then replicated the mixed-effects logistic regression from the previous section separately on each experimental half, which confirmed that differences among demands were stable (Figure 6, right). Together, the consistency and stability displayed by participants suggest that decision makers had a clear representation of the environment.

@@

  Understanding the sporadic attractiveness of effortful options remains elusive. In this study, we tested the hypothesis that such appeal could arise from being able to compare different means to achieve the same outcome. Using two foraging experiments that exposed individuals to a single or intermixed types of demands, we found that 1) preferences for cognitive effort were stable when individuals face a single form of demand, but fade over time when other costs are also experienced; and 2) the phenomenon can be explained by a model in which preferences depend on how participants construe the amount of time a task takes, which could hypothetically be recalibrated when all tasks are interleaved.

@@

>>>>

---
title: Apparent preferences for cognitive effort fade when multiple forms of effort
  and delay are interleaved in a foraging environment
author: "Claudio Toro-Serey^1^2, Gary Kane^1^2, & Joseph T. McGuire^1^2"
header-includes:
- \usepackage{setspace}\doublespacing
- \usepackage[left]{lineno}
- \linenumbers
- \usepackage{float}
- \newcommand{\beginsupplement}{ \setcounter{table}{0} \renewcommand{\thetable}{S\arabic{table}}
  \setcounter{figure}{0} \renewcommand{\thefigure}{S\arabic{figure}} }
- \usepackage{indentfirst}
output:
  pdf_document:
    highlight: haddock
    keep_tex: no
  word_document: #default
    reference_docx: Style_reference_draft.docx
  html_document:
    df_print: paged
csl: apa.csl
indent: true
#subtitle: 'Abbreviated Title: Heterogeneity in the functional topography of mPFC'
bibliography: cost_paper.bib
---

@@

1.  Department of Psychological and Brain Sciences, Boston University, Boston, USA

2.  Center for Systems Neuroscience, Boston University, Boston, USA

@@

  Evaluating whether to pursue a desired outcome often requires assessing the demands required for obtaining it. In psychology and economics, demands for both time and effort have traditionally been identified as costs that detract from the net subjective value of rewards [@Shenhav2017 ;@Westbrook2015]. Indeed, decision makers tend to prefer low-effort and immediate rewards over larger rewards that demand longer delays and higher effort [@Ainslie1975; @Hull1943; @Kool2010; @Kool2018; @Treadway2009; @Westbrook2013]. However, people often change their attitudes towards exerting effort in everyday decisions, resulting in pursual of effortful avenues instead. Emerging experimental and theoretical work has chiefly explained this phenomenon by suggesting a reciprocal relationship between effort and outcome, whereby exerting effort boosts the perceived value of its outcomes [@HernandezLallement2014; @Kacelnik2002; @Mochon2012], and positive outcomes imbue value into their preceding effortful behavior [i.e. learned industriousness; @Eisenberger1992]. Though intuitively valid and empirically supported, this explanation leaves open questions about what features of an effortful task can make it attractive in its own right, and how contextual factors mold its perceived costs [@Inzlicht2018]. Understanding these intrinsic qualities could improve our ability to motivate completion of effortful daily activities, such as schoolwork or physical exercise.

  What is often missing from existing accounts is the common case of choosing among types of behavioral demands. Studies have traditionally compared the reduction of subjective value imposed by each type of demand, often through discounting functions that capture how much less appealing a reward becomes as a function of the amount or intensity of its associated behavior [@Ainslie1975; @Frederick2002; @Green1994; @Kable2007; @Kool2010; @Westbrook2013]. Although this line of work suggests that discounting functions are similar for delay and effort [@Massar2015; @Prevost2010; @Seaman2018], there is ongoing debate about the shape of the discounting function across effort types [@Arulpragasam2018; @Chong2017; @Klein-Flugge2016; @Kool2018; @Prevost2010], and discount rates do not always correlate across cost domains [@Seaman2018; @Westbrook2013]. However, sporadic preferences for effort could arise from the availability of other types of demands in the same choice environment—in other words, from its context [@Kool2018]. For example, exerting high effort for a reward might be aversive, but the same level of work might become valuable when compared to the boredom of passive waiting (e.g. taking the cumbersome way home rather than waiting for a delayed train, even if the resulting time of arrival is the same). Conversely, even though people tend to donate more to charitable events that require them to perform effortful activities (i.e. the martyrdom effect), they are also less likely to engage in such events once an easier alternative to raise funds becomes available [@Olivola2013]. Recent work has proposed that the subjective costs produced by delays and cognitive effort is modulated by the possibility of missing out on better concurrent alternatives, namely the opportunity cost of time [OC; @Fawcett2012; @Kurzban2013; @Otto2019]. This evidence sets the stage for examining what people would actually select when given the option for multiple types of demands, including the unexplored question of deciding between cognitive and physical effort [@Inzlicht2018; @Kool2018].

  Single choice foraging paradigms provide a natural way to probe preferences among demands, by asking individuals whether the current offer outweighs what could be obtained and experienced during that time instead [@Stephens2019]. Foraging has recently attracted high scientific interest due to its ecological validity, which is rooted in evolutionary choice behaviors [@Hayden2018; @Mobbs2018], and offers a number of desirable qualities. First, it allows for the estimation of optimal, reward maximizing choice patterns based on the opportunity cost of time [@Charnov1976; @Krebs1977]. We can therefore interpret observed deviations in acceptance rates from such optimal behavior in addition to comparing them among demands. Second, the sequential nature of foraging allows individuals to emphasize different elements of the decision context. For example, the cost (or value) of effort or delay could globally depend on the availability of alternative demands (as suggested above), impacting the perceived opportunity cost of time. Alternatively, it could be intrinsic and robust to such manipulation, and/or depend on task-specific factors that are adjusted once individuals experience different forms of behavioral costs. Foraging rests on an abundant modeling history that can help to computationally capture such emphases [@Stephens2019].

@@

  All experimental tasks were implemented using PsychoPy 2 [v1.85.1, @Peirce2019] on a Macbook Pro laptop. In this task, participants foraged for monetary rewards in an environment in which each trial required either physical effort or cognitive effort for a set period of time (the “handling time”), or an equivalent unfilled delay (Figure 1). Their goal was to maximize their gains within a fixed amount of time. On each trial a monetary offer was displayed for 2 s, and participants had the opportunity to expend time and/or effort during the handling time in order to earn it. Upon completion of a trial participants saw a 2 s window displaying the reward obtained, which was followed by a travel time to the next offer. Alternatively, the participant could quit at any point during the handling time by pressing the spacebar on the computer, and immediately start traveling.

  Each participant was assigned to one of three types of costs (cognitive effort, physical effort, or delay), and a fourth group of equal size faced an effortless physical task that involved minimal gripping. Each group was unaware that other cost conditions existed. Participants exerted physical effort by maintaining grip on a handheld dynamometer (*VERSION*, BIOPAC Systems, United States) using their dominant hand. Gripping requirements were calibrated at 20% of maximum voluntary contraction (MVC, acquired at the beginning of the session). Cognitive effort entailed switching among Stroop, dot motion coherence, and flanker tasks. In Stroop, one of three color names was displayed on the screen (red, blue or green), with a font color that was either congruent (e.g. word red painted in red) or incongruent (e.g. word red painted in blue). Participants had to select the color of the font, not the word displayed (i.e. they had to suppress their tendency to read the word). For motion coherence, 100 solid white dots moved on the screen. A fraction of these dots moved cohesively to the left or right, while the rest moved in random directions (coherence could be either 30% or 40%, uniformly sampled). Participants had to respond with the direction of the cohesive set of dots. In flanker, rows of arrowheads pointed either to the left or the right (maximum of 3 rows, 3 to 13 arrowheads per row). Participants responded with the direction of the center arrowhead, which could point in the same or opposite direction from its neighbors. These tasks were configured so that responses always involved a left or right key press (e.g. for Stroop, two colored circles were presented at each side of the screen). During the handling time, cognitive tasks and their configurations were randomly sampled, and were presented for 1 s followed by a 1 s inter-stimulus interval. Participants were asked to respond within each task’s presentation time. Before the experiment, participants trained in each cognitive task until they correctly performed six consecutive tasks of each kind. While in the handling time, if participants failed to maintain above-threshold gripping or made two mistakes during the cognitive task, they were forced to travel and missed the reward.

  There were three block types, in which handling times of 2, 10, or 14 seconds were paired with travel times of 14, 6, and 2 seconds, respectively (note that all combinations add up to 16 seconds). Timing parameters were held constant within each 7-min block. Each of these blocks was experienced in a pseudo-randomized order, and repeated in the same order after a short break in order to probe choice stability (total session length = 42 minutes). Reward amounts varied uniformly per trial (4, 8, or 20 cents), with the constraint that every reward was presented twice every six trials. This prevented sequences from being dominated by a single amount during any window of time. Timing information was disclosed at the beginning of each block, and rewards displayed during a 2 s offer window before each trial began. Participants received training prior to the experimental session, and were told about all possible environmental statistics in order to preclude biases due to experience-dependent learning [@Dundon2020; @Garrett2020].

@@

  Foraging theory posits that accepting a delayed reward should depend on the opportunity cost (OC) of time incurred in obtaining it, given by the richness of the environment [@Charnov1976; @Stephens2019]. In this study, the richness of the environment was manipulated by the length of the handling and travel times, where shorter travel times produced richer environments. Since time combinations were fixed per block, we calculated each block’s optimal accept/reject strategy by computing all decision strategies according to the following equation:

@@

  All analyses were performed in R 3.5.2 [@RTeam2018]. First we tested whether decision makers integrated delay and reward information. To address the prediction that participants would be more likely to accept higher reward and shorter handling time trials, we fit a mixed-effects logistic regression to predict trial-wise acceptances [using the lme4 package, @Bates2015], giving a random intercept to each subject to account for biases in participant behavior. We included regressors for handling time and reward, as well as two covariates that probed the influence of recent history on choices. The first was an autoregressive regressor containing the number of consecutive offer misses prior to a given trial (misses could be due to quits or forced travels). The second regressor tracked the sum of the previous *n* offers. We identified the *n* that minimized Akaike's Information Criterion (AIC) among 6 versions of the model (tracing from 2 previous trials up to 7, beyond which the model failed to converge), and reported all coefficients from the winning model.

  Our foraging task was configured such that over and under accepting were detrimental to total earnings. To confirm this, we fit a general linear model with constant, linear, and quadratic terms to estimate the correspondence between proportion of trials completed (independent variable) and total earnings (dependent variable). A significant quadratic coefficient thus would signal that the task statistics operated as expected. Next, to determine the optimality of each group’s decisions, we performed two-sided one-sample t-tests to see if the proportion of acceptances for each time/reward combination was significantly different from the optimal rate (see Operationalization of Cost). This resulted in 36 independent tests (3 rewards amounts, 3 handling/travel time combinations, and 4 groups), so we corrected for multiple comparisons using False Discovery Rate [FDR, @Benjamini2001].

@@

  We started by fitting a base model derived from the normative predictions of the Marginal Value Theorem [MVT; @Charnov1976], which states that a delayed reward should be accepted if its earning rate surpasses the opportunity cost (OC) of time incurred in obtaining it. This decision rule can be expressed as

@@

$$\begin{aligned}
P(accept)_i = \dfrac {1} {1 + e^{-(1/\beta)(Model_{base})}}
\end{aligned}$$

@@

  A final model considered the relationship between effortful engagement and the perceived length of time. Being immersed in a task can make time transpire subjectively faster [@Csikszentmihalyi2014], which might have made the cognitive task attractive. Time sensitivity has been previously used to improve fits on classic hyperbolic models of delay discounting [@Mckerchar2009]. It is possible that such subjective handling time was recalibrated in the within-subjects experiment, resulting in the observed convergance of acceptances across demands. We therefore added a time sensitivity parameter $S$ to the model, with a separate value associated with each type of demand. The values of $S_{demand}$ could then be recalibrated as participants experienced them, averaged according to the following formula [adapted from @Sutton2018]

@@

  in which high values of the free parmeter $\alpha_w$ denoted that decision makers incorporated the recalibrated subjective time early on in the experiment. Similar to @Mckerchar2009, the values of $mS_i$ exponentially modulated every instance of the handling time ($H_i$) from the adaptive model. In this way, subjective time perception produced by different demands could influence choice in two ways: retrospectively affecting the value of $\gamma$ through the perceived length of recent trials

@@

  The next set of analyses examined the hypothesis that effort group would uniformly accept more trials. Figure 3 (left) shows that the cognitive effort group consistently accepted more offers, which resulted in lower earnings (middle). One-way ANOVAs showed that differences among groups were significant both for overall acceptance rates (F(`r propAov[[1]]$Df[1]`, `r propAov[[1]]$Df[2]`) = `r round(propAov[[1]][1, 4], digits = 2)`, *p* `r report_p(propAov[[1]][1, 5])`), and total earnings (F(`r earnAov[[1]]$Df[1]`, `r earnAov[[1]]$Df[2]`) = `r round(earnAov[[1]][1, 4], digits = 2)`, *p* `r report_p(earnAov[[1]][1, 5])`). Post-hoc permutations (5000 iterations) comparing mean acceptance rates between every pair of costs confirmed that acceptance rates of the cognitive effort group (mean = `r filter(overallSumms, Cost == "Cognitive")$mProp`, SD = `r filter(overallSumms, Cost == "Cognitive")$sdProp`) were higher than those of the physical (mean = `r filter(overallSumms, Cost == "Physical")$mProp`, SD = `r filter(overallSumms, Cost == "Physical")$sdProp`; *p* `r report_p(filter(propTests, contrast == "Physical < Cognitive")$permPval)`, Cohen's D = `r filter(propTests, contrast == "Physical < Cognitive")$cohenD`) and wait groups (mean = `r filter(overallSumms, Cost == "Wait")$mProp`, SD = `r filter(overallSumms, Cost == "Wait")$sdProp`; *p* `r report_p(filter(propTests, contrast == "Cognitive > Wait")$permPval)`, Cohen's D = `r filter(propTests, contrast == "Cognitive > Wait")$cohenD`). It also showed that those faced with the easy task accepted more than those in the wait group, although responses from the easy group were more variable (mean Easy = `r filter(overallSumms, Cost == "Easy")$mProp`, SD = `r filter(overallSumms, Cost == "Easy")$sdProp`; *p* `r report_p(filter(propTests, contrast == "Wait < Easy")$permPval)`, Cohen's D = `r filter(propTests, contrast == "Wait < Easy")$cohenD`) (all other *p* > 0.05). The higher acceptance rates produced by the cognitive effort group resulted in significantly lower earnings (mean = `r filter(overallSumms, Cost == "Cognitive")$mEarn`, SD = `r filter(overallSumms, Cost == "Cognitive")$sdEarn`) than both physical (mean = `r filter(overallSumms, Cost == "Physical")$mEarn`, SD = `r filter(overallSumms, Cost == "Physical")$sdEarn`; *p* `r report_p(filter(earnTests, contrast == "Physical > Cognitive")$permPval)`, Cohen's D = `r filter(earnTests, contrast == "Physical > Cognitive")$cohenD`) and wait groups (mean = `r filter(overallSumms, Cost == "Wait")$mEarn`, SD = `r filter(overallSumms, Cost == "Wait")$sdEarn`; *p* `r report_p(filter(earnTests, contrast == "Cognitive < Wait")$permPval)`, Cohen's D = `r filter(earnTests, contrast == "Cognitive < Wait")$cohenD`). In addition, we found that the wait group earned more than the physical (*p* `r report_p(filter(earnTests, contrast == "Physical < Wait")$permPval)`, Cohen's D = `r filter(earnTests, contrast == "Physical < Wait")$cohenD`) and easy groups (mean Easy = `r filter(overallSumms, Cost == "Easy")$mEarn`, SD = `r filter(overallSumms, Cost == "Easy")$sdEarn`; *p* `r report_p(filter(earnTests, contrast == "Wait > Easy")$permPval)`, Cohen's D = `r filter(earnTests, contrast == "Wait > Easy")$cohenD`). The percentage of forced travels for cognitive effort was `r round(propFails[2, 2] * 100, digits = 2)`% across participants, and 0% for either physical or easy conditions, making performance an unlikely explanation for the observed differences in choice.

@@

  Stability was defined as a persistence in acceptance patterns throughout the experimental session. The scatter plot on Figure 5 (bottom-left) shows the overall proportion accepted by each participant before and after the mid-point break. Paired permutations showed no differences in acceptances for wait and easy groups (*p* > 0.05), but we found significantly lower acceptances for physical (*p* `r report_p(prepost_btw$Perms[1])`, Cohen's D = `r round(prepost_btw$CohenD[1], digits = 2)`) and cognitive (*p* `r report_p(prepost_btw$Perms[2])`, Cohen's D = `r round(prepost_btw$CohenD[2], digits = 2)`) groups. Figure 6 (left) shows that the reduction acceptances in effort groups over time was steeper for longer handling times, a hallmark of effortful demands [@Treadway2009]. We then replicated the mixed-effects logistic regression from the previous section separately on each experimental half, which confirmed that differences among demands were stable (Figure 6, right). Together, the consistency and stability displayed by participants suggest that decision makers had a clear representation of the environment.

@@

  Understanding the sporadic attractiveness of effortful options remains elusive. In this study, we tested the hypothesis that such appeal could depend on the ability to achieve the same outcome through different means (**not quite**). Using two foraging experiments that exposed individuals to a single or intermixed types of demands, we found that 1) preferences for cognitive effort were stable when individuals face a single form of demand, but fade over time when other costs are also experienced; and 2) the phenomenon can be explained by a model in which preferences depend on how participants construe the amount of time a task takes, which could hypothetically be recalibrated when all tasks are interleaved.

  Our findings provided partial support for our pre-registered hypotheses, which can be found in https://osf.io/2rsgm/. In accord with our predictions, decision makers were successful in approximating reward-maximizing strategies, as evidenced by acceptance rates. This is in line with previous foraging experiments, where the main deviation from optimality is a tendency to overharvest. Unlike our first hypothesis, our predictions regarding differences among costs did not

  Confirming our predictions, we found that individuals could approximate optimality by preferring higher rewards and shorter delays, regardless of the richness of the environment or whether they faced a single or intermixed demands. When faced with a single type of demand, people in the cognitive effort group accepted more trials than optimal, whereas physical effort and wait groups under-accepted, partially matching our predictions and highlighting that cognitive effort can add value to outcomes. However, against our expectations, we found that a similar preference for cognitive effort disappeared over time as decision makers foraged for types of demands in addition to rewards. Computational modeling suggested that cognitive effort was valuable because it shortened the subjective elapsed time of trials, a perception that could be recalibrated as individuals experienced multiple forms of demands. The shorter time sensitivity of cognitive effort could be due to its constant recruitment of working memory, impeding estimations of the nominal time, or because of the discrete nature of the task (versus a continuous configuration of the physical and wait tasks). These results expand our understanding of the attractiveness of effort, bringing forth outcome-independent features that can be leveraged to motivate effort engagement in diverse everyday scenarios.

@@

- AFAIK, Eisenberger1982 provides evidence that task diversity motivates further effort engagement

## References
